### 2020-02-29 doc/icfp20/intro.tex
wip

Full dependent types allow a fully expressive logic, and it is
tempting to take a standard textbook on music theory such as
\citet{piston-harmony} or \citet{aldwell2018harmony} and attempt to
formalize it in type theory.

-- people are starting to use computers for analysis and synthesis of rule systems
(reference Haskell work). In particular FP, and some rudimentary dep types

-- With full dep types, expressive logic, tempting to formalize all the rules
of some theory book in dep types. Point out difficulties....

-- Music Tools, compare to Euterpea + other Haskell work, mention FARM.

-- In this paper, focus on counterpoint plus basic start to some harmony.

Counterpoint section: contrast old and new methods

Harmony section: Harmonize melody, use counterpoint for voice leading and to contrain harmony.
Mention grid for harmonic analsis, makes use of vectors.

conclusion: easy to express rules, break into parts, reuse in other contexts, extend.
Downsides: hard to do some things (equality, sets, etc), Proof emphasis can get in the way, but
may want to use in the future. Future work: equivalence, ornaments, ML.

TODO: Fix bib style.

>>>>

Full dependent types allow expression of full predicate logic, and it is
tempting to take a standard textbook on music theory such as
\citet{piston-harmony} or \citet{aldwell2018harmony} and attempt to
formalize it in type theory. However this is difficult since most
``rules'' in music theory are not absolute, but rather suggestions of
varying degrees (themselves vaguely specified) of importance. Since
ultimately the resulting created music is more important than the
theory, an alternative is to attempt to indirectly deduce the rules
from music samples, using for example machine
learning~\citep{huang-cp}. However.... TODO
TODO: Perhaps create rules and weight them, then use ML.

Although formalization of a large amount of music theory is a goal,
for now we start with the modest task of expressing a small,
relatively strict rule set known as species
counterpoint~\cite{fux-cp}. TODO--more here
We also present preliminary work on harmonizing a melody using
a subset of rules based on \citet{piston-harmony}, contrasting with
existing work by \citet{koops-fharm}. Notably, since counterpoint and
harmony are not separate concepts but in fact deeply intertwined, we
are able to reuse counterpoint rules to help develop natural-sounding
harmonizations.

We make use of and extend~\cite{MusicTools}, a library of small tools
for analysis and synthesis of music written in Agda, with the goal
of eventually being a dependently-typed alternative to Euterpea. 
Previous work by ~\cite{cong-cp} using Music Tools expressed
first-species counterpoint with the Agda type system, so that an error
in counterpoint resulted in an Agda type error. Although this takes
advantage of Agda's native type checking, it has several
downsides. One is that the Agda errors may be difficult to interpret
as the corresponding musical error. Another is TODO: flexibility and
composability.

This is an experience report, and we highlight both the advantages and
disadvantages of using Agda

TODO: check that bib style is allowed.
